---
title: "Building footprint area packing"
author: "Iva Brunec"
format: html
editor_options: 
  chunk_output_type: inline
---

## Load the data

Read in the data from the [EUBUCCO](https://eubucco.com/data/) website (this takes a relatively long time):

```{r}
#| label: read-data
#| echo: true
#| message: false
#| cache: true
library(sf)
dat_full <- read_sf('v0_1-SVN.gpkg')

```

This dataset is > 1 mil rows, for the purposes of this visualization, I'm only taking 1,000 randomly sampled rows

```{r}
dat <- dat_full[sample(nrow(dat_full), 1000), ]  
```

## Calculate area of each polygon

Each row in the 'geom' column contains a polygon. We can use an sf package function to calculate the area of each polygon and save it in a separate column.
For plotting, we'll assume each polygon is a square (which it isn't), and save the sides of the square as well.


```{r}
#| label: calculate-area
#| echo: true
#| message: false
library(dplyr)
library(knitr)
dat$area <- st_area(dat$geom) 

dat <- dat |>
  mutate(side_x = sqrt(area), side_y = sqrt(area)) |>
  st_drop_geometry() |>
  # save only coordinates to pass on to python
  select(side_x, side_y)

dat$side_x <- as.numeric(dat$side_x)
dat$side_y <- as.numeric(dat$side_y)

kable(head(dat))

```

## Rectangle packing

Now, we pass these data to python & use the `rpack` library to derive coordinates to pack rectangles into a plot:

```{python}

dat_py = r.dat
# round to nearest integer, otherwise the rpack function doesn't work
dat_py = round(dat_py, 0).astype(int)

# convert to list of arrays
dat_list = dat_py.values.tolist()

import rpack
import pandas as pd
import numpy as np

positions = rpack.pack(dat_list)

positions = pd.DataFrame(np.row_stack(positions))
positions.columns =['x_coord', 'y_coord']

positions = pd.concat([positions.reset_index(drop=True), dat_py], axis=1)

```

## Visualization

Port the python results back to R and visualize.

```{r}
#| label: viz
#| echo: true
#| message: false
#| fig-align: center
#| fig-alt: "Rectangle packing representation of building footprints"
library(reticulate)
library(ggplot2)
library(showtext)
library(ggtext)
data_coord <- py$positions

# combine the two dataframes
#data_final <- cbind(dat, data_coord)

data_coord$x_end = data_coord$x_coord + data_coord$side_x
data_coord$y_end = data_coord$y_coord + data_coord$side_y
# this is slightly stupid, but whatever. recalculate the area
data_coord$area <- data_coord$side_x * data_coord$side_y

color_scale <- c(
  "#271220", 
  "#24101e", 
  "#574557", 
  "#a192ad", 
  "#e0d3f3", 
  "#e0c8e5"
)

font_add_google(name = 'Major Mono Display', family = 'MajorMono')
showtext_auto()

ggplot() +
  geom_rect(data = data_coord, 
            aes(xmin = x_coord, ymin = y_coord,
                xmax = x_end, ymax = y_end,
                fill = log(area)), color = '#F2eef8',
            linewidth = .08) +
  scale_fill_gradientn(colours=rev(color_scale)) +
  labs(title = '1,000 BUILDING FOOTPRINTS',
       caption = 'Each square represents the area of one of 1,000 randomly sampled building footprints in slovenia. <br>Plotted with rectangle-packer. Data: eubucco.\t') +
  coord_fixed() +
  theme_void() +
  theme(legend.position = 'none',
        plot.background = element_rect(fill = '#F2eef8', color = NA),
        plot.title = element_text(size = 130, family = 'MajorMono',
                                  color = '#271220', hjust = 0.5),
          plot.caption = element_markdown(lineheight = .4,
                                          size = 36, 
                                          family = 'MajorMono',
                                          color = '#271220'))


ggsave('day_29_monochrome.png', width = 12, height = 8, dpi = 300)

```
